Sizing the bets in a focused portfolio

### 2023-10-03 doc/paper.tex
Minor updates to the paper

The answer to this question has been given by Kelly with his widely known Kelly
formula (sometimes called Kelly criterion)~\cite{kelly}. The Kelly's approach
starts by maximizing long--term growth of capital when one is presented with
an infinite amount of opportunities to bet on. This work extends that idea by
considering:
\begin{enumerate}
    \item Multiple companies (stocks) in parallel,
    \item An arbitrary number of scenarios for each company,
    \item Multiple constraints for modelling our preference towards no shorting,
    no use of leverage, and providing a maximum value for the risk of permanent
    loss of capital,
    \item A fundamentals--based analysis (after all, a stock is an ownership
    share of business~\cite{intelligentInvestor}).
\end{enumerate}

@@

The underlying philosophy is that one should spend the majority of his time
analyzing investments and thinking about intrinsic values under different
scenarios that might play out, independently of outside thoughts and events.
However, calculating intrinsic value of a company is more of an art than a
science, especially for a high-quality, growing businesses within one's circle
of competence. And according to Charlie Munger, Warren Buffett, Mohnish Pabrai
and the like-minded others from whom the authors got the inspiration for this
work, one should focus precisely on getting such great businesses for a fair
price. That means that one shouldn't take what this approach says at face value,
and one should probably use its guidance infrequently.

@@

In addition to the most important margin of safety mentioned above, there are
however a couple of more margins of safety embedded in the current framework.
These are:
\begin{itemize}
    \item No shorting allowed. From a purely mathematical point of view,
    shorting would be allowed. Without detailed math, it is easy to see how a
    company with a negative expected return would result in a short position.
    However, due to the asymmetry of the potential losses compared to gains,
    coupled with the usual time--frame limit that comes with shorting, we
    provided a constraint for being long--only.
    \item No use of leverage allowed. Again, from a purely mathematical point of
    view, use of leverage would sometimes be useful. The thinking in avoiding to
    use leverage is that no--one should be in a hurry to get rich, and should
    avoid risking good night's sleep based on short--term market fluctuations,
    which are fairly hard to predict consistently (unless you work at
    Renaissance Technologies, in which case you can scratch this).
    \item No company without at least one downside scenario is allowed. By
    disallowing inputs without downside, the framework forces you to focus and
    think about what can go wrong, as opposed to dreaming about what can go
    right. However, if one is absolutely (100\%) sure that a company does not
    have a downside, the solution is to put all the money in that one company.
    It is our feeling that the best way to handle such cases is outside of this
    framework. Alternatively, one can always model an unknown downside scenario
    with a small probability (say 5--20\%) and intrinsic value of zero.
\end{itemize}

@@

\noindent If one assumes an infinite number of (re)allocations
$N_a$\footnote{The assumption regarding the infinite number of allocations is
something that the authors are slightly uncomfortable with, but \textit{feels}
it is fine because of the margins of safety embedded into the thinking that goes
into assessing each investment opportunity.}, the following relation holds:
\begin{equation}
\label{eq:7}
    \lim_{N_a \to \infty} \frac{n_i}{N} = p_i
\end{equation}

@@

\noindent With the basic validation of the numerical model performed, the
attention is moved to a realistic example. Consider five candidate companies,
each with up to three scenarios. Each scenario is represented by an intrinsic
value and the probability of the scenario happening (or intrinsic value being
reached at some point in the future). Note that how these numbers are obtained
is outside of the scope of this work, although it is important to stress that
the validity and conservative assumptions behind these numbers are probably the
most important part of an investor's job. The example inputs are presented
in~\autoref{tab:companyA} to~\autoref{tab:companyE}.

@@

\noindent The last item is particularly interesting to the authors. According to
Actuarial Life Tables in~\cite{lifeTables}, the probability of the (currently 34
year--old, the oldest) author dying within the next year is approximately
0.26\%\footnote{This is the only thing that's probably worth measuring in the
basis points.}. That is two orders of magnitude higher than the probability of
the permanent loss of capital for this portfolio. Considering that the portfolio
has five stocks, that is a very strong argument against excessive
diversification, especially if:
\begin{itemize}
    \item One thinks of stocks as ownership shares of businesses, which implies
    long--term thinking and not being bothered by market fluctuations, 
    \item One embeds a margin of safety in different scenarios for different
    companies by e.g. recognizing that both unknown and known bad things may
    happen.
\end{itemize}

@@

>>>>

The answer to this question has been given by Kelly with his widely known Kelly
formula (sometimes called Kelly criterion)~\cite{kelly}. The Kelly's approach
starts by maximizing long--term growth of capital when one is presented with
an infinite amount of opportunities to bet on. This work extends that idea by
considering:
\begin{enumerate}
    \item Multiple companies (stocks) in parallel,
    \item An arbitrary number of scenarios for each company,
    \item Multiple constraints for modelling our preference towards no shorting,
    no use of leverage, and providing a maximum value for the risk of permanent
    loss of capital,
    \item A fundamentals--based analysis with a very long time horizon. After
    all, a stock is an ownership share of business~\cite{intelligentInvestor},
    in which time--scales are significantly larger than the time--scales of
    day--to--day stock price fluctuations.
\end{enumerate}

@@

The underlying philosophy is that one should spend the majority of their time
analyzing investments and thinking about intrinsic values under different
scenarios that might play out, independently of outside thoughts and events.
However, calculating intrinsic value of a company is more of an art than a
science, especially for a high-quality, growing businesses within one's circle
of competence. And according to Charlie Munger, Warren Buffett, Mohnish Pabrai
and the like-minded others from whom the authors got the inspiration for this
work, one should focus precisely on getting such great businesses for a fair
price. That means that one shouldn't take what this approach says at face value,
and one should probably use its guidance infrequently.

@@

In addition to the most important margin of safety mentioned above, there are
however a couple of more margins of safety embedded in the current framework.
These are:
\begin{itemize}
    \item No shorting allowed. From a purely mathematical point of view,
    shorting would be allowed. Without detailed math, it is easy to see how a
    company with a negative expected return would result in a short position.
    However, due to the asymmetry of the potential losses compared to gains,
    coupled with the usual time--frame limit that comes with shorting, we
    provided a constraint for being long--only.
    \item No use of leverage allowed. Again, from a purely mathematical point of
    view, use of leverage would sometimes be useful. The thinking in avoiding to
    use leverage is that no--one should be in a hurry to get rich, and should
    avoid risking good night's sleep based on short--term market fluctuations,
    which are fairly hard to predict consistently (unless you work at
    Renaissance Technologies, in which case you can scratch this).
    \item No company without at least one downside scenario is allowed. By
    disallowing inputs without downside, the framework forces you to focus and
    think about what can go wrong, as opposed to dreaming about what can go
    right. This is in--line with thoughts by Prasad~\cite{darwinInvesting} about
    the importance of avoiding the errors of investing in a bad company.
    However, if one is absolutely (100\%) sure that a company does not have a
    downside, the solution is to put all the money in that one company.  It is
    our feeling that the best way to handle such cases is outside of this
    framework. Alternatively, one can always model an unknown downside scenario
    with a small probability (say 5--20\%) and intrinsic value of zero.
\end{itemize}

@@

\noindent If one assumes an infinite number of (re)allocations
$N_a$\footnote{The assumption regarding the infinite number of allocations is
something that the authors are slightly uncomfortable with, but \textit{feel}
it is fine because of the margins of safety embedded into the thinking that goes
into assessing each investment opportunity.}, the following relation holds:
\begin{equation}
\label{eq:7}
    \lim_{N_a \to \infty} \frac{n_i}{N} = p_i
\end{equation}

@@

\noindent With the basic validation of the numerical model performed, the
attention is moved to a realistic example. Consider five candidate companies,
each with up to three scenarios. Each scenario is represented by an intrinsic
value and the probability of the scenario happening (or intrinsic value being
reached at some point in the future). Note that how these numbers are obtained
is outside of the scope of this work, although it is important to stress that
the validity and conservative assumptions behind these numbers are probably the
most important part of an investor's job who is taking a fundamental analysis
approach. The example inputs are presented in~\autoref{tab:companyA}
to~\autoref{tab:companyE}.

@@

\noindent The last item is particularly interesting to the authors. According to
Actuarial Life Tables in~\cite{lifeTables}, the probability of the (currently 34
year--old, the oldest) author dying within the next year is approximately
0.26\%\footnote{This is the only thing that's probably worth measuring in the
basis points in this framework.}. That is two orders of magnitude higher than
the probability of the permanent loss of capital for this portfolio. Considering
that the portfolio has five stocks, that is a very strong argument against
excessive diversification, especially if:
\begin{itemize}
    \item One thinks of stocks as ownership shares of businesses, which implies
    long--term thinking and not being bothered by market fluctuations, 
    \item One embeds a margin of safety in different scenarios for different
    companies by e.g. recognizing that both unknown and known bad things may
    happen.
\end{itemize}

@@

\noindent The software is freely available at TODO. Reader is referred to its
documentation for instructions on how to install and use the software.
